---
title: "Project Report"
author: "Devanshi, Chinmay, Archie"
date: "4/10/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
library (readr)
```
### INTRODUCTION AND DATA
##### Background and Motivation
The American police force is notorious for needlessly beating and killing. Moreover, such killings and mistreatment tend to disproportionately impact Black Americans and other poc groups. In May of 2020, the issue again sparked outrage, as the nation watched policeman Derek Chauvin kill George Floyd on video. The event brought discussion on the impacts of police brutality, particularly on communities of color, to a national spotlight. Our research study will analyze which communities and groups of people are most impacted by police killings. 

##### Dataset in question
The data overviews the demographics of individuals killed by police in the United States. This data was collected from The Guardian’s database on police killings, and census data from the American Community Survey. The Guardian’s database data was collected through verified crowdsourcing. The data was collected in 2015, and was used in the FiveThirtyEight article “Where Police Have Killed Americans in 2015.” The data is posted to the github page “fivethirtyeight/data” and can be found by the link. Relevant variables in the dataset include raceethnicity : Race/ethnicity of deceased, state : State where incident occurred, share_white : share of population that is nonhispanic / white, share_bloack: share of population that is black, month: month of killing.

### Research Question and Hypothesis

Research Question: Among those who were killed by police, are the household incomes of White individuals greater than Black & Hispanic individuals?

Null Hypothesis: The household incomes of White individuals who were killed by police are not greater than Black & Hispanic individuals who were killed by police

Alternate Hypothesis: The household incomes of White individuals who were killed by police are greater than Black & Hispanic individuals who were killed by police

dataset: https://github.com/fivethirtyeight/data/tree/master/police-killings 
background reading: https://www.nytimes.com/2020/05/31/us/george-floyd-investigation.html 

### Glimpse of Dataset:

```{r}
urlfile="https://raw.githubusercontent.com/fivethirtyeight/data/master/police-killings/police_killings.csv"
police_killings<-read_csv(url(urlfile))
glimpse (police_killings)
```

### METHODOLOGY
```{r}
police_killings <- police_killings %>%
  mutate(black_hispanic = case_when(
    raceethnicity == "Black" ~ "B/H",
    raceethnicity == "Hispanic/Latino" ~ "B/H",
    raceethnicity == "White" ~ "W"
  )) 

p_hat <- police_killings %>%
  specify(h_income ~ black_hispanic) %>%
  calculate(stat = "diff in means", order = c("W", "B/H")) 
  

killings_boot <- police_killings %>%
  specify(h_income ~ black_hispanic) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("W", "B/H")) 
percentile_ci <- get_ci(killings_boot)

visualize(killings_boot) +
  shade_confidence_interval(endpoints = percentile_ci) + 
  labs(title = "Race/Ethnicity Difference in Means Distribution for Household Income", 
  x = "Race/Ethnicity", y = "Household Income")

p_hat
percentile_ci
```

Using sample data to make conclusions about the underlying population the sample came from is statistical inference. In our research, we want to examine whether household incomes of all White individuals in the U.S. killed by police are greater than that of Black & Hispanic individuals killed with a sample data of size 467, so we have to use statistical inference. To be more specific, we uses functions from the "infer" package to create 10,0000 bootstrap simulation, calculate bootstrap stats "differences in means" , and construct a 95% confidence interval of difference of mean household income between white and non-white victims (h_income_white - h_income_B/H). Then we visualize our confidence interval and bootstrap distribution using visualize() & shade_confidence_interval() function. If the confidence interval captures zero on the plot, we fail to reject our null hypothesis; if not, we have convincing evidence that household incomes of white killed by police is not the same as that of Black & Hispanic killed.

### RESULTS
We obtain a 95% confidence interval from 3346	to 11023, meaning that we are 95% confident that the interval [3346, 11023] captures the true difference of mean household income (white - black & hispanic). The interval is consistently positive, meaning that we are confident to say that the mean household income of white killed by police is greater than that of black or hispanic killed by police, at alpha = 0.05, based on our bootstrap simulation model.

